%----------------------------------------------------------------------------------------------------
% START OF CHAPTER
%----------------------------------------------------------------------------------------------------
\chapter{Apr 27}
\paragraph{Topics}
\textit{Intro to the past tense. Sentences with the past tense. Grammar analysis.}

\section{Grammar} \label{0427-Grammar}
\subsection{Past tense - \iti{passato prossimo}}

In English, `I have spoken' and `I spoke' are grammatically equivalent in everyday use. In Italian,
everyday speech and writing uses the latter form, with \iti{essere} and \iti{avere} acting as helper verbs.
The past tense therefore uses a three-word structure: \textit{subject, helper-verb, past-participle}.

Broadly speaking, transitive verbs use \iti{avere} and intransitive verbs and verbs of motion use \iti{essere}.
See \cite{SanFilippo} and \cite{Ragusa} for more explanation.

\paragraph{Verbs that require \itb{essere}}
\begin{itemize}[nosep]
    \item \iti{essere}
    \item \iti{stare}
    \item \iti{entrare}
    \item \iti{arrivare}
    \item \iti{venire}
    \item \iti{andare}
    \item \iti{uscire}
    \item \iti{partire}
\end{itemize}


To form the past participle, remove the ending of the verb (e.g., \iti{-are, -ere, -ire} and then add 
\iti{-ato, -uto, -ito} to match. If \iti{essere} is the helper verb, then change the ending to match gender and plurality. 

\begin{itemize}[nosep]
    \item \iti{comprare} $ \rightarrow $ \iti{comprato}
    \item \iti{lavorare} $ \rightarrow $ \iti{lavorato}
    \item \iti{vedere}   $ \rightarrow $ \iti{visto}    
    \item \iti{arrivare} $ \rightarrow $ \iti{arrivato} or \iti{arrivata, arrivati} 
\end{itemize}

\begin{table}[ht]
\centering
    \begin{tabular}{ r | l }
                       & Helper: \itb{avere} Verb: \itb{comprare}    \\
        \hline
        \itb{io}       & \iti{ho comprato}      \\
        \itb{tu}       & \iti{hai comprato}     \\
        \itb{lui/lei}  & \iti{ha comprato}      \\
        \itb{noi}      & \iti{abbiamo comprato} \\
        \itb{voi}      & \iti{avete comprato}   \\
        \itb{loro}     & \iti{hanno comprato}  
    \end{tabular}
    \caption{Past tense example: \iti{avere + comprare}}
    \label{tab:tPastTenseAvereComprare}
\end{table}

\begin{table}[ht]
\centering
    \begin{tabular}{ r | l }
                 & Helper: \itb{essere} Verb: \itb{andate}    \\
         \hline
        io       & \iti{sono andato}   \\
        tu       & \iti{sei andato}    \\
        lui      & \iti{è andato}      \\
        lei      & \iti{è andata}      \\
        noi      & \iti{siamo andati}  \\
        voi      & \iti{siete andati}  \\
        loro     & \iti{sono andati}  
    \end{tabular}
    \caption{Past tense example: \iti{essere + andare}}
    \label{tab:tPastTenseEssereAndare}
\end{table}

Unsurprisingly, some verbs are irregular and you just have to learn them.

\begin{table}[ht]
\centering
    \begin{tabular}{ r | l }
         Past participle   & Helper: \itb{avere} Irregular verbs   \\
         \hline
        \textbf{said, told}     & \iti{detto}   \\
        \textbf{made, did}      & \iti{fatto}   \\
        \textbf{read}           & \iti{letto}   \\
        \textbf{seen, saw}      & \iti{visto}   \\
        \textbf{put}            & \iti{messo}   \\
        \textbf{wrote, written} & \iti{scritto}  
      \end{tabular}
    \caption{Irregular past participles: all with \iti{avere}}
    \label{tab:tIrregularPastParticiples}
\end{table}

\paragraph{Examples}
\begin{itemize}
  \item I'm happy because my wife made a really great dish of pasta last night. We watched a good move after dinner. \\
  \iti{Io sono felice perché mia moglie ha hatto un piatto di pasta molto buono ieri sera. 
    Noi abbimamo vitso un buon film dopo cena.}

  \item I said that I worked a lot today and am tired. I don't want to go to the restaurant for 
    dinner. I prefer to stay at home. \\
  \iti{Ho detto che ho lavorato molto oggi e sono stanco. Io non voglio andare al ristorante per cena. 
    Preferisco stare a casa.}
\end{itemize}

\section{Grammar analysis - sentence corrections}
\begin{itemize}
  \item Test: \iti{Io non sono comprando una casa in Italia perché è molto cara.} \\
  Should be: \iti{Io non sto comprando una casa in Italia perché è molto cara.} \\
  \textit{Correction: use \iti{stare} with present participle.} \\
  
  \item Test: \iti{Io non posso parla adesso perché io sto lavorando.} \\
  Should be: \iti{Io non posso parlare adesso perché io sto lavorando.} \\
  \textit{Correction: use infinitive \iti{parlare} with helper verbs.} \\
  	
  \item Test: \iti{Mio libro è molto difficile.} \\
  Should be: \iti{Il mio libro è molto difficile.} \\
  \textit{Correction: Need article with possessives except for family.} \\

  \item Test: \iti{Io mangio questo piatto di pasta perché è molto buono.} \\
  Should be: \iti{Io mangio questa piatto di pasta perché è molto buono.} \\
  \textit{Correction: Need to use \iti{questa} to match gender of \iti{pasta}, not \iti{piatto}.}
\end{itemize}
%----------------------------------------------------------------------------------------------------
% END OF CHAPTER
%----------------------------------------------------------------------------------------------------
