%----------------------------------------------------------------------------------------------------
% START OF CHAPTER
%----------------------------------------------------------------------------------------------------
\chapter{Feb 7}
\paragraph{Topics}
\textit{Grammar recap, More about adjectives, Possessive adjectives revisited}
\section{Grammar} \label{0207-Grammar}
\subsection{Review}
\begin{itemize}[nosep]
    \item \itb{essere}: \iti{Io sono, tu sei, lui/lei è, noi siamo, voi siete, loro sono}
    \item \itb{avere}: \iti{Io ho, tu hai, lui/lei ha, noi abbiamo, voi avete, loro hanno}
    \item \itb{stare}: \iti{Io sto, tu stai, lui/lei sta, noi stiamo, voi state, loro stanno}
\end{itemize}
\paragraph{Examples}
\begin{itemize}
    \item  You are happy because you stay at your home. \\
    \iti{Tu sei felice perché tu stai a tua casa.}
    \item  I am studying Italian today. \\
    \iti{Io sto studiando oggi.}
    \item  I am happy because I have a new car. \\
    \iti{Io sono felice perché io ho una nuova macchina.}\footnote{\iti{Macchina} is more common and less formal than \iti{auto}, like `car' vs. `automobile' or `vehicle'.}
    \item  I am happy because I have the car. \\
    \iti{Io sono felice perché io ho la macchina.}
    \item  The car is beautiful. \\
    \iti{La macchina è bella.}
    \item  I am talking now. \\
    \iti{Io sto parlando adesso.}
    \item  Claudio is driving now. \\
    \iti{Claudio sta guidando adesso.}
    \item  I am not happy because I am not eating pizza. \\
    \iti{Io non sono felice perché io non sto mangiando la pizza.}
    \item  You are sad because you do not have a new car. \\
    \iti{Tu sei triste perché tu non hai una nuova macchina / auto.}
    \item  I am good. \\
    \iti{Io sono bravo.} (performance)
    \item  I am a good dad. \\
    \iti{Io sono un buon papà.\footnote{Accents matter \ldots \iti{Papà} is `dad' but \iti{Papa} is `Pope'.}}
    \item  I am good (when you are asked the question ``how are you''). \\
    \iti{Io sto bene.}
    \item  I am not happy because my wife is not cooking pizza this evening / tonight. \\
    \iti{Io non sono felice perché mia moglie non sta cucinando la pizza questa sera.}
    \item  This book is a good book. \\
    \iti{Questo libro è un buon libro.}
\end{itemize}
\subsection{Adjectives - \iti{buono / bello / quello}}
The adjectives \iti{buono}, \iti{bello}, and \iti{quello} (`good', `beautiful', and `that', respectively) have a different spelling when they are used before a masculine noun. See Table \ref{tab:tAdjectivesOfQualityBeforeMascNouns}. The adjectives are not shortened for feminine nouns.

\begin{table}[ht]
\centering
    \begin{tabular}{ r | r l }
         \textbf{base form} & \textbf{becomes...} & \textit{example}   \\
         \hline
         buono   & buon  & \iti{Un buon gelato}     \\
         bello   & bel   & \iti{Un bel telefono}    \\
         quello  & quel  & \iti{Quel tavolo}
    \end{tabular}
    \caption{\iti{Buono}, \iti{bello}, and \iti{quello} adjectives preceding masculine nouns}
    \label{tab:tAdjectivesOfQualityBeforeMascNouns}
\end{table}
\paragraph{Examples}
\begin{itemize}
    \item Good day! \\
    \iti{Buon giorno!}
    \item It's a good gelato. \\
    \iti{È un buon gelato.}
    \item That gelato is good. \\
    \iti{Quel gelato è buono.}
    \item It's a good pizza. \\
    \iti{È una buona pizza.}
    \item This pizza is good. \\
    \iti{Questa pizza è buona.}
\end{itemize}

\subsection{Possessive adjectives}
Saying, ``The my car is red'' is wrong in English. Italian on the other hand uses this construction, with the articles, a lot, and especially with possessive adjectives.

Possessive adjectives change ending to match the number and gender of the noun that they relate to, just like other adjectives. For example, \iti{la sua macchina} can mean `his car' or `her car', because the \iti{sua} matches the noun \iti{macchina.}

\begin{table}[ht]
\centering
    \begin{tabular}{ r | r | l }
         \textbf{adjective} & \textbf{meaning} & \textit{example} \\
         \hline
         \itb{mio}   & my, mine  & \iti{Il mio gelato}      \\
         \itb{mia}   &           & \iti{La mia macchina}    \\ 
         \itb{tuo}   & your \textit{sing.}  & \iti{Il tuo gelato}    \\
         \itb{tua}   & your \textit{sing.}  & \iti{La tua macchina}  \\
         \itb{suo}   & his/her   & \iti{Il suo gelato}      \\
         \itb{sua}   &           & \iti{La sua macchina}    \\
         
         \itb{nostro} & our      & \iti{Il nostro vino}     \\
         \itb{nostra} & our      & \iti{La nostra macchina} \\
    
         \itb{vostro} & your \textit{pl.}\tablefootnote{Plural as in ``you all's'' or ``y'all's''} & \iti{Il vostro vino}     \\
         \itb{vostra} & your \textit{pl.} & \iti{La vostra macchina} \\
    \end{tabular}
    \caption{Possessive adjectives}
    \label{tab:tAdjectivesPossessive}
\end{table}
\begin{mdframed}
  Remember that when we talk about family members (humans, not the dog!) we can drop the article, e.g., \iti{mia moglie} instead of \iti{la mia moglie}.    
\end{mdframed}


\paragraph{Examples}
\begin{itemize}
    \item My car is red. \\
    \iti{La mia macchina è rossa.}
    \item His car is red. \\
    \iti{La sua macchina è rossa.}
    \item Our son is very happy because he has a new car. \\
    \iti{Nostro figlio è molto felice perché lui ha una nuova macchina.}
    \item My dog is very intelligent. \\
    \iti{Il mio cane è molto intelligente.}
    \item My red wine is very good. \\
    \iti{Il mio vino rosso è molto buono.}
    \item Your wine is very good. (Talking to a group.)\\
    \iti{Il vostro vino è molto buono.}
    \item Your wine is very good. (Talking to an individual.) \\
    \iti{Il tuo vino è molto buono.}
    \item My husband is not cooking pasta tonight. \\
    \iti{Mio marito non sta cucinando la pasta questa sera.}
\end{itemize}

\section{Vocabulary} \label{0207-Vocabulary}
\subsection{Verbs - \iti{lavorare} and other \iti{-are} examples}

\begin{table}[ht]
\centering
    \begin{tabular}{ r l | r l | r l | r l }
         \textbf{verb}     &          &           &                   &             &                 &                  &                  \\
         \hline
         \itb{lavorare} & to work  & \itb{io} & \iti{lavoro} & \itb{tu} & \iti{lavori} & \itb{lui/lei} & \iti{lavora}  \\
         \itb{riposare} & to rest  &             & \iti{riposo} &             & \iti{riposi} &                  & \iti{riposa}  \\
         \itb{studiare} & to study &             & \iti{studio} &             & \iti{studi}  &                  & \iti{studia}  \\
         \itb{cucinare} & to cook  &             & \iti{cucino} &             & \iti{cucini} &                  & \iti{cucina}  \\
         \itb{mangiare} & to eat   &             & \iti{mangio} &             & \iti{mangi}  &                  & \iti{mangia}  \\
         \itb{parlare}  & to speak &             & \iti{parlo}  &             & \iti{parli}  &                  & \iti{parla}   
    \end{tabular}
    \caption{Verbs: \iti{-are} forms}
    \label{tab:tVerbs-Are}
\end{table}

\section{Homework} \label{0207-Homework}
\begin{enumerate}
    \item I am busy. I am working now. \\
    \iti{Io sono occupato. Io sto lavorando adesso.}
    
    \item I am not well. I am resting. \\
    \iti{Io non sto bene. Io sto riposando.}
    
    \item I am sick (adjective). I am not working tonight. \\
    \iti{Io sto male. Io non sto lavorando questa sera.}
    
    \item I am not sick. I am working today. \\
    \iti{Io non sto male. Io lavoro oggi.}
    
    \item My son is studying Italian. \\
    \iti{Mio figlio sta studiando italiano.}
    
    \item My car is not fast. \\
    \iti{La mia macchina non è rapida.}
    
    \item My wine is very good. \\
    \iti{Il mio vino è molto buono.}
    
    \item I am very happy because my wife is cooking pizza tonight. \\
    \iti{Io sono molto felice perché mia moglie sta cucinando la pizza questa sera.}
    
    \item I am not happy because I am not eating pizza tonight. \\
    \iti{Io non sono felice perché io no sto mangiando la pizza questa sera.}
    
    \item She is very happy because she is not working today. \\
    \iti{Lei è molto felice perché lei non lavora oggi.}
    
    \item This pizza is very good. You are very good (performance). \\
    \iti{Questa pizza è molta buona. Tu sei molto bravo!}
    
    \item Your wife does not speak Italian. I speak Italian. \\
    \iti{Tua moglie non parla italiano. Io parlo italiano.}
\end{enumerate}
    
%----------------------------------------------------------------------------------------------------
% END OF CHAPTER
%----------------------------------------------------------------------------------------------------
