%----------------------------------------------------------------------------------------------------
% START OF CHAPTER
%----------------------------------------------------------------------------------------------------
\chapter{Mar 28}
\paragraph{Topics} 
\textit{Irregular verbs: \iti{fare} and \iti{andare}. Imperative tense. Pronunciation tips, directions, need. Homework review.}

\section{Grammar}
\subsection{Irregular verbs: \iti{fare} and \iti{andare}}
Some verbs you just have to know. Two important ones are \iti{fare}, to do/to make, and \iti{andare}, to go.
Fare is used in lots of idiomatic expressions. See Table \ref{tab:tVerbFareAndAndare}.

\begin{table}[ht] %[ht] % place table roughly [h]ere or else at the [t]op of page
	\centering
	\begin{tabular}{ r | l |  l  }
		              & \itb{fare}     &  \itb{andare}   \\
		\hline              
		\itb{io}      & \iti{faccio}   &  \iti{vado}     \\
		\itb{tu}      & \iti{fai}      &  \iti{vai}      \\
		\itb{lui/lei} & \iti{fa}       &  \iti{va}       \\
		\itb{noi}     & \iti{facciamo} &  \iti{andiamo}  \\
		\\
		\textit{gerund} & \iti{facendo} & \iti{andando}
	\end{tabular}
	\caption{Some restaurant terms}
	\label{tab:tVerbFareAndAndare}
\end{table}

\paragraph{Examples}
\begin{itemize}
	\item{I'm taking a shower.}   \\
	  \iti{Io faccio la doccia}
	\item{What do you do for work?}\footnote{Culturally, asking about your job is not really done in Italy.}  \\
	  \iti{Cosa fai per lavoro?}
	\item{I'm going to the store.}   \\
      \iti{Io sto andando in negozio.}
	\item{I need to go to the store.}   \\
      \iti{Io ho bisogno di andare al negozio.}
	\item{Are you going on vacation?}   \\
      \iti{Tu vai in vacanza?} or \iti{Tu stai andando in vacanza?}
\end{itemize}

\subsection{The imperative tense}
This verb form is used to offer suggestions or commands from the speaker to a listener. 
Consequently, it only exists in the second person and first person plural.
For example, in Table \ref{tab:tVerbImperative}, the commands would be ``Speak!'' and ``Let's talk!''  

\begin{table}[ht] %[ht] % place table roughly [h]ere or else at the [t]op of page
	\centering
	\begin{tabular}{ r | l |  l  }
		              & \itb{present}  &  \itb{imperative} \\
		\hline              
		\itb{io}      & \iti{parlo}    &                   \\
		\itb{tu}      & \iti{parli}    &  \iti{parla}      \\
		\itb{lui/lei} & \iti{parla}    &                   \\
		\itb{noi}     & \iti{parliamo} &  \iti{parliamo}   \\
		\itb{voi}     & \iti{parlate}  &  \iti{parlate}    
	\end{tabular}
	\caption{Comparison of present and imperative verb tense, with \iti{parlare}}
	\label{tab:tVerbImperative}
\end{table}

\paragraph{Examples}
\begin{itemize}
	\item Eat! Eat!        \\
	  \iti{Mangia, mangia!}
	\item Let's go!        \\
	  \iti{Andiamo!}
	\item Drive!           \\
	  \iti{Guida!}
	\item Listen!          \\
	  \iti{Ascolta!}
	\item Wait!            \\
	  \iti{Aspetta!}
	\item Get in! (e.g., a car) \\
	  \iti{Entra!} 
\end{itemize}

\section{Words}
\subsection{Directions}
When you ask for directions, e.g., by asking, ``\iti{Dov'è il Colosseo?}'', you are 
likely to get back a sequence of instructions in the imperative tense: `turn left at the 
lights', `go straight', etc. The same works in Italian.

\begin{table}[ht] %[ht] % place table roughly [h]ere or else at the [t]op of page
	\centering
	\begin{tabular}{ r | l | r l }
		\itb{infinitive}  &  \itb{imperative} & Example               &  \\
		\hline
		\itb{andare}      & \iti{vai}         & \iti{Vai diritto}     & Go straight \\
		\itb{girare}      & \iti{gira}        & \iti{Gira a destra}   & Turn right \\
		\itb{svoltare}    & \iti{svolta}      & \iti{Al semaforo, svolta a sinistra} & At the lights, turn left \\
		\itb{tornare}     & \iti{torna}       & \iti{Torna indietro}  & Turn around \\
		\itb{passare}     & \iti{passa}       & \iti{Passa la chiesa} & Pass the church
	\end{tabular}
	\caption{Some imperatives, used in giving directions}
	\label{tab:tDirectionsImperative}
\end{table}

Some other useful words for directions (\ref{tab:tVocabDirections}):

\begin{table}[ht] %[ht] % place table roughly [h]ere or else at the [t]op of page
	\centering
	\begin{tabular}{ r | l }
		left           &  \iti{a sinistra}  \\
		right          &  \iti{a destra}    \\
		near           &  \iti{vicino}      \\
		far            &  \iti{lontano}     \\
		after          &  \iti{dopo}        \\
		intersection   &  \iti{incrocio}    \\
		traffic light  &  \iti{semaforo}    \\
        stop (transit) &  \iti{fermata}
	\end{tabular}
	\caption{Some common terms, used in giving directions}
	\label{tab:tVocabDirections}
\end{table}

\paragraph{Examples}
\begin{itemize}
	\item At the intersection, turn left. \\
	  \iti{Al semaforo, gira a sinistra.}
	\item This is my stop. \\
      \iti{Questa è la mia fermata.}
	\item The church is not far away. \\
      \iti{La chiesa non è lontana.}
	\item At the next intersection, turn right, then left.  \\
      \iti{Al prossimo incrocio, gira a destra, e poi a sinistra.}
	\item Turn around at the bank, and go straight ahead. \\
	  \iti{Torna indietro al banca, e vai diritto.}  
\end{itemize}

\subsection{\itb{avere bisogno di} - to have need of}
Use the three word combination of \iti{avere}, \iti{bisogno}, and \iti{di} 
to express `to have a need of' something. 

\begin{itemize}
	\item I need a coffee. \\
	\iti{Io ho bisogno di un caffè.}
\end{itemize}

\section{Homework}
\begin{enumerate}
	\item{I would drive but I am very tired.}        \\
	  \iti{Io guiderei, ma sono molto stanco.}
	\item{I want to buy the wine at the restaurant.} \\
	  \iti{Io voglio comprare il vino al ristorante.}
	\item{I can’t talk now. I am working.}           \\
	  \iti{Io non posso parlare adesso. Io sto lavorando.}
	\item{I would like to order, please.}            \\
	  \iti{Vorrei ordinare, per favore.}
	\item{More wine, please.}                        \\
	  \iti{Più vino, per favore.}
	\item{Nice to meet you. My name is (your name).} \\
	  \iti{Piacere. Il mio nome è Shriman.}
	\item{Hello. I would like to buy a bottle of wine for my friend.} \\
	  \iti{Salve. Vorrei comprare una bottiglia di vino per il mio amico.}
	\item{Where is a bank?}                \\
	  \iti{Dov'è una banca?}
	\item{I am fine, thank you. And you?}   \\
	  \iti{Io sto bene, grazie. E tu?}
	\item{I am on vacation for one week.}  \\
	  \iti{Io sono in vacanza per un settimana.}
	\item{The wine is in the car.}         \\
	  \iti{Il vino è nella macchina.}
	\item{Is there a discount?}            \\
	  \iti{C'è uno sconto?}
	\item{I am studying italian in Virginia.}           \\
	  \iti{Io sto studiando italiano in Virginia.}
	\item{I can’t. I am at work now.}                   \\
	  \iti{Non posso. Io sono a lavoro adesso.}
	\item{Hello. We have a reservation for two.}        \\
	  \iti{Salve. Noi abbiamo una prenotazione per due.}
	\item{The pizza in this restaurant is very good.}   \\ 
	  \iti{La pizza in questo ristorante è molto buona.}
	\item{I would like the bill, please. We are ready.} \\ 
	  \iti{Vorrei il conto, per favore. Siamo pronti.}
	\item{My friend does not speak Italian. I speak a little Italian.} \\
	  \iti{Il mio amico non parla italiano. Io parlo un po' d'italiano}
	\item{I would try this dish. Thank you.}             \\
	  \iti{Io proverei questo piatto. Grazie.}
	\item{Can I have a bottle of natural water, please?} \\
	  \iti{Posso avere una bottiglia d'acqua naturale, per favore?}
\end{enumerate}

%----------------------------------------------------------------------------------------------------
% END OF CHAPTER
%----------------------------------------------------------------------------------------------------
