%----------------------------------------------------------------------------------------------------
% START OF CHAPTER
%----------------------------------------------------------------------------------------------------
\chapter{Apr 18}
\paragraph{Topics} 
\textit{Grammar analysis, Adverbs of place: `here' and `there'. Sentence review.}

\section{Grammar} \label{0418-Grammar}
\subsection{Grammatical sentence review}
\begin{itemize}
  \item I am talking with Claudio.  \\
    \iti{Io sto parlando con Claudio.}
  \item I am tired and I would like to go home.  \\
    \iti{Io sono stanco e io vorrei andare a casa.}
  \item I am sick and I would like to stay home. (Female speaker)  \\
    \iti{Io sono malata (io sto male) e io vorrei stare a casa.}
  \item I am happy because I am going to Italy.  \\
    \iti{Io sono felice perché io sto andando in Italia.}
  \item I can go, but I am very tired tonight.  \\
    \iti{Io posso andare, ma, io sono molto stanco questa sera.}
  \item I would eat pizza tonight because I do not have time to cook.  \\
    \iti{Io mangerei pizza questa sera (stasera) perché io no ho tempo di cucinare.}
  \item I am not driving to go to Rome because I like the train.  \\
    \iti{Io non sto guidando per andare a Roma perché mi piace il treno.}
  \item My family is in Sicily, and I would like to go there.  \\
    \iti{La mia famiglia è in Sicilia, e io vorrei andare là.}
  \item I am not traveling because the flight ticket is very expensive.  \\
    \iti{Io non sto viaggiando perché il biglietto di volo è molto caro.}
 \end{itemize}
 
 \subsection{Here and there - \iti{qui}, \iti{qua}, \iti{lì}, and \iti{là}}
 English speakers use `here' and `there' to refer to both generalized and specific locations, with 
 stress making the distinction. `I left my keys here.' versus `I left my keys \textit{here}.'
 In Italian, different adverbs are used, as shown in \ref{tab:HereAndThereQuiQuaLiLa}.
 \footnote{Mnemonic: the ones ending \iti{-a} are `approximate´, the ones ending \iti{-i} are `i\sout{e}xact'.}
 Also see \cite{Broster}.
 
 \begin{table}[ht] %[ht] % place table roughly [h]ere or else at the [t]op of page
 	\centering
 	\begin{tabular}{ r | l | l }
 		\textbf{here}           & \iti{qua}    & \textit{general vicinity}  \\
 		\textbf{there}          & \iti{là}     & \textit{general vicinity}  \\
 		                        &              &                            \\
 		\textbf{\textit{here}}  & \iti{qui}    & \textit{specific location} \\
        \textbf{\textit{there}} & \iti{lì}     & \textit{specific location} \\
 	\end{tabular}
 	\caption{Location adverbs - here and there}
 	\label{tab:HereAndThereQuiQuaLiLa}
 \end{table}
 
\section{Sentence review}
\begin{enumerate}
	\item I'm happy because the pizza is good. \\ 
	  \iti{Io sono felice perché la pizza è buona.}
	\item My wine is good.  \\ 
	  \iti{Il mio vino è buono.}
    \item My brother is in Italy.  \\ 
      \iti{Mio fratello è in Italia.}
    \item I'm studying with Claudio's book.  \\
      \iti{Io studio con il libro di Claudio.}
    \item The book is on the table. \\
      \iti{Il libro è sul tavolo.}
    \item{I'm going home.} \\
      \iti{Io sto andando a casa.}    
    \item I am happy. \\
      \iti{Io sono felice.}
    \item I have the pizza. \\
      \iti{Io ho la pizza.}
    \item I eat the pizza. \\
      \iti{Io mangio la pizza.}
    \item I am well / unwell. \\
      \iti{Io sto bene / male.}        
    \item I am talking with Claudio.  \\ 
      \iti{Io sto parlando con Claudio.}
    \item I would eat pasta.  \\
      \iti{Io mangerei pasta.}
    \item I would like the pasta. \\
      \iti{Io vorrei la pasta}
    \item I can eat pizza.  \\
      \iti{Io posso mangiare pizza.}
    \item I have to eat pizza.  \\
      \iti{Io devo mangiare pizza.}
    \item I want to eat pizza.  \\
      \iti{Io voglio mangiare pizza.}
\end{enumerate}

%----------------------------------------------------------------------------------------------------
% END OF CHAPTER
%----------------------------------------------------------------------------------------------------